Lichat, A Lightweight Chat Protocol

### 2022-01-24 els2022-lichat/paper.tex
Sigh

\section{Introduction}\label{introduction}
The IRC (Internet Relay Chat) protocol has long stood the test of time as a reliable protocol for text chat communication. Large networks and a plethora of mature clients still exist and are in use today. However, the IRC protocol is also plagued by a number of issues that stem from its historical design and largely distributed evolution. In particular, the protocol has many different, incompatible extensions, and features many idiosyncrasies that make it difficult to write new clients or servers.

@@

\section{Related Work}\label{relatedwork}
IRC

Matrix

XMPP

Whatever?

@@

\section{Further Work}\label{further-work}
On the protocol side, we would like to develop further extensions, primarily to allow for VOIP, federation, and encrypted messaging.

@@

\section{Acknowledgements}\label{acknowledgements}
We would like to thank all the very sexy people that read this paper ;)

>>>>

\section{Introduction}\label{introduction}
The IRC\cite{oikarinen1993internet} (Internet Relay Chat) protocol has long stood the test of time as a reliable protocol for text chat communication. Large networks and a plethora of mature clients still exist and are in use today. However, the IRC protocol is also plagued by a number of issues that stem from its historical design and largely distributed evolution. In particular, the protocol has many different, incompatible extensions, and features many idiosyncrasies that make it difficult to write new clients or servers.

@@

\section{Related Work}\label{relatedwork}
Penpoppichanan et al.\cite{penpoppichanan2001security} identify a number of security and privacy related issues with the IRC protocol. While Lichat shares the basic concerns about privacy, being an unencrypted protocol, it does fix several of the identity and authentication related issues identified in the thesis.

The XMPP core specification\cite{saint2004extensible} alone is over 200 pages long, and does not include the already non-trivial question of parsing XML, which is used for its wire format. Saint-Andre et al.\cite{saint2009xmpp} also point out the lack of a proper end-to-end encryption scheme for the protocol.

Jennings et al.\cite{jennings2006study} compare a variety of Instant Messenger systems such as AIM and MSN, which are all now largely defunct. The ppaper focuses primarily on the distribution of sessions and the scalability of the networks. Lichat is explicitly focused on small to mid-sized communities and is thus not concerned with the scalability to tens of thousands of simultaneous users.

@@

\section{Further Work}\label{further-work}
On the protocol side, we would like to develop further extensions, primarily to allow for VOIP, federation, and encrypted messaging. Encryption and privacy are of especially big interest, though also pose quite challenging as end-to-end encrypted one-to-many communication is difficult to achieve efficiently\cite{nabeel2017many}\cite{whisper2014}.

@@

\section{Acknowledgements}\label{acknowledgements}
We would like to thank Georgiy Tugai, Jan Moringe, and Selwyn Simsek for their feedback on this paper and contributions to Lichat's development.
